---
title: "p8105_hw1_rcn2115"
author: "Riya Nadkarni"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

# Problem 1

#Loading the dataset

```{r}
data("penguins", package = "palmerpenguins")
```
# Description of the dataset

This dataset contains the measurements and characteristics of penguins and includes variables like **species**, **island,bill_length_mm**, **bill_depth_mm**, **flipper_length_mm**, **body_mass_g**, **sex** and **year**.

This dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns. 

The mean flipper length is `r round(mean(penguins$flipper_length_mm, na.rm=TRUE),2)` mm

# Creating a scattterplot

```{r, yx_scatter}
ggplot(data = penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + 
  geom_point()
ggsave("flipper_vs_bill_scatterplot.png")
```

# Problem 2

#Creating a dataframe

```{r, df_create}
hw1_df = 
  tibble(
    norm_samp=rnorm(10),
    vec_logical = norm_samp > 0,
    vec_char = c("A","B","C","D","E","F","G","H","I","J"),
    vec_factor= factor(c("low","medium","high","low","medium","high","low","medium","high","low"))
  )
print(hw1_df)
```

#Taking the mean of each variable in the dataframe

This works and gives the mean of the numeric vector.
```{r}
mean(pull(hw1_df, norm_samp))
```

This works since the logical vector is treated as numeric (TRUE = 1 and FALSE = 0). 
```{r}
mean(pull(hw1_df, vec_logical))
```

This does not work as characters cannot be averaged. 
```{r}
mean(pull(hw1_df, vec_char))
```

This does not work as factors need to be converted to numeric before their mean can be calculated.
```{r}
mean(pull(hw1_df, vec_factor))
```

# Converting logical, character, and factor variables to numeric
```{r, eval = FALSE}
converted_vec_logical = as.numeric(hw1_df$vec_logical)
converted_vec_char = as.numeric(hw1_df$vec_char)
converted_vec_factor = as.numeric(hw1_df$vec_factor)
```

The **logical** values are converted to numerical values (1 and 0). TRUE becomes 1 and FALSE becomes 0. This explains why we get the mean for these values. 

The **character** variable conversion gives a warning ( Warning: NAs introduced by coercion). This explains why we get "NA" when we try to take the mean.

The **factor** levels (non-numeric) are assigned integral values. This explains why we can get the mean for these values. 